---
title: "Hypothesis Testing II"
author: "Arash Hatamirad, ahatamirad@gmail.com"
date: "9/1/2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = "C:/Users/Sanaz/Documents/All_MSDA_Projects and Sources/R Ptojects/Algorithm I/HM01/Algorithm-I-HW01")

# Install required packages

#install.packages("MASS")
#install.packages("fBasics")
#install.packages("moments")
#install.packages("dplyr")


library(MASS)
library(fBasics)
library(moments)
library(dplyr)
```

The airquality data will be used for this Exercise. The information about variables in the dataset can be found in R.

## Exercise: Hypothesis Testing  

Perform a hypothesis test -whether Wind in July has a different speed (mph) than Wind in August.    

#### Question (a): 
- Which test should we perform, and why? See QQ-plot and perform Shapiro-Wilk test for normality check.  


#### Answer -:  
- we must compare "Wind" variable in July with "Wind" variable in August  
- Since we have "Two-Populations", we have to check normality for both samples.  

```{r}

# Filter the July & August data

wind.july=airquality %>% 
  filter(airquality$Month==7)

wind.august=airquality %>% 
  filter(airquality$Month==8)


```

##### Visual method (QQ Plot)  
 

```{r}

# divide 1X2 
par(mfrow=c(1,2))

# QQ Plot for wind, July
qqnorm(wind.july$Wind,main = "QQ Plot for Wind - July",ylab="MPH")
qqline(wind.july$Wind,col="red")


# QQ Plot for wind, August
qqnorm(wind.august$Wind,main = "QQ Plot for Wind - August",ylab = "MPH")
qqline(wind.august$Wind,col="red")



```

##### Quantitave method   
##### Shapiro-Wilk Test 

- H0: Data follows normal distribution  
- H1: Data doesn't follow normal distribution  

```{r}
# shapiro-Wilk Test

#Print P-Value for July
wind.july$Wind %>% 
  shapiro.test() %>% 
  print()

# Print P-Value for August
wind.august$Wind %>% 
  shapiro.test() %>% 
  print()


```
 
##### comments: 

According to the QQ-Plot, there are a majority points near the red line and regarding to P-Values:  
 P-Value > Significant Value  
therefore, we don't have enough evidence to reject the "Null Hypothesis", hence both data follow normal distribution.  


##### Conclusion: 
Both data follow normal distribution, therefore we should perform "T-Test".  


#### Question (b): 
- Specify null and alternative hypotheses


#### Answer -:  

- Null Hypothesis (H0):  
 - Mean of wind in July = Mean of wind in August, it means Wind in July has similar speed (mph) as Wind in August.  

- Alternate Hypothesis (H1):  
 - Mean of wind in July != Mean of wind in August, it means Wind in July has different speed (mph) with Wind in August.  


#### Question (c): 
- State the conclusion based on the test result.  


#### Answer -: 
1- We must check the variances for both data is equal or not.  

- H0 : Two groups have the same variances  
- H1 : Two groups have different variances  

```{r}
# Make a Union data frame, incluse July and August data
wind.july.august=union(wind.july,wind.august)

# Check the Variances
var.test(wind.july.august$Wind ~ wind.july.august$Month, alternative="two.sided")

```

##### comments: 

According to the P-Value, because it's very bigger than significant value, thus we don't have enough evidence to reject H0    

- Two groups have the same variances, hence  we will run Pooled T-Test  

```{r}
# Run pooled T-Test

t.test(wind.july.august$Wind ~ wind.july.august$Month,alternative="two.sided",var.equal=TRUE)


```

#### Conclusion: 
- P-Value >> significant value  
P-Value is very bigger then significant value,therefor we don't have enough evidence to reject H0   

- result: Mean of wind in July = Mean of wind in August, it means Wind in July has similar speed (mph) as Wind in August.    


***